ProteInfer: deep networks for protein functional inference

### 2021-09-20 latex/the_supplement.tex
Updates

\subsubsection*{Label inheritance}
Some annotations, such as GO terms and EC numbers, come as directed acyclic graphics (which take the form simple trees for EC numbers). Typically in such cases the annotation provided on UniProt is the most-specific that is known. For example if a protein is known to exhibit \textit{sequence specific DNA binding} (GO:0043565). It will not separately be annotated with the ancestral term \textit{DNA binding} (GO:0003677), this is simply assumed from the ontology. Using such an annotations directly, however, is likely to be problematic in a deep learning setting. Failing to annotate an example with the parental term demands that the model predict that the example is negative for this term, which is not the effect we want.
To address this our datasets include labels for all ancestors of applied labels for EC, GO and InterPro datasets. In the case of GO we restrict these to \textit{is\_a} relationships.

\subsubsection*{Class activation mapping}
Many proteins have multiple functional properties. For example we analyse the case of the bifunctional dhfr/ts of \textit{T. gondii}. Such bifunctional enzymes are often not unique -- it is functionally advantageous for these enzymes to be fused, which facilitates channeling of substrate between their active sites. Since there are a number of such examples within Swiss-Prot, the mere existence of a TS domain in a protein is (mild) evidence for possible DHFR function. To increase the interpretability of the network we introduce a technique that simplifies such cases by subtracting the activations for other predicted classes during class activation mapping.

>>>>

\subsubsection*{Label inheritance}
Some annotations, such as GO terms and EC numbers, are structured as directed acyclic graphics (which take the form of simple trees for EC numbers). Typically in such cases the annotation provided on UniProt is the most-specific that is known. For example if a protein is known to exhibit \textit{sequence specific DNA binding} (GO:0043565). It will not separately be annotated with the ancestral term \textit{DNA binding} (GO:0003677), this is simply assumed from the ontology. Using such an annotations directly, however, is likely to be problematic in a deep learning setting. Failing to annotate an example with the parental term demands that the model predict that the example is negative for this term, which is not the effect we want.
To address this our datasets include labels for all ancestors of applied labels for EC, GO and InterPro datasets. In the case of GO we restrict these to \textit{is\_a} relationships.

\subsubsection*{Class activation mapping}
Many proteins have multiple functional properties. For example we analyse the case of the bifunctional dhfr/ts of \textit{T. gondii}. Such bifunctional enzymes are often not unique -- it is functionally advantageous for these enzymes to be fused, which facilitates channeling of substrate between their active sites. Since there are a number of such examples within Swiss-Prot, the mere existence of a TS domain in a protein is (mild) evidence for possible DHFR function. To increase the interpretability of the network we subtract the activations for other predicted classes from those of the class of interest during class activation mapping.
